Adaptive evolution with aneuploidy and mutation

### 2021-10-25 ms/ms.tex
updated prior text for fitnesses

\paragraph{Prior distributions.} The prior distributions of $w_{2n+1}=1-c+b$, $w_{2n+1^*}=(1+s)(1-c)+b$ and $w_{2n^*}=1+s$ were obtained by estimating $1-c$, $1-c+b$ and $1+s$ from growth curves data. These growth curve data were previously obtained by~\citet{Yona2012} using mono-culture growth experiments.
The raw growth curves data were not published before, but were used to produce figures 3C, 4A, and S2 in \citet{Yona2012}.
We used \texttt{Curveball}\footnote{https://curveball.yoavram.com}, a dedicated method for predicting results of competition experiments from growth curve data~\citep{Ram2019}. \texttt{Curveball} takes growth curves of two strains growing separately in mono-culture and predicts how they would grow in a mixed culture, that is, it predicts the results of a competition assay.
From these predictions, relative fitness values can be computed.  
We used growth curves of $2n+1$ and $2n$ in \SI{39}{\celsius} and in \SI{30}{\celsius} to estimate $1-c+b$ and $1-c$, respectively. The latter assumes that the cost of aneuploidy is the same in \SI{39}{\celsius} and \SI{30}{\celsius}; this might be incorrect, but we only assume this to generate a prior distribution for the fitness values. We also used growth curves of $2n+1$ and $2n^*$ in \SI{39}{\celsius} to estimate $1+s$. 
Because \texttt{Curveball} uses a maximum-likelihood approach to estimate model parameters, we were able to estimate a distribution of relative fitness values by sampling from a truncated multivariate normal distribution defined by the maximum-likelihood covariance matrix. Thus, we sampled 10,000 values for $1-c+b$, $1-c$, and $1+s$, which we used as prior distributions. Note that the first and last values correspond to $w_{2n+1}$ and $w_{2n}$, whereas $w_{2n+1^*}=(1-c)(w_{2n^*}-1)+w_{2n+1}$.
See Figures~\ref{fig:growth-curves-30deg} and~\ref{fig:growth-curves-39deg}.
%TODO ILIA please report here the point estimate and 95% CI for the three distributions, I only have for 1-c+b: 0.9863 (95% CI 0.9790 - 0.9939).
%TODO ILIA describe the priors for the mutation and aneuploidy rate.
% TODO also describe the prior for the model without aneuploidy - the mutation rate is wider here.

@@

>>>>

\paragraph{Prior distributions.} The prior distributions of $w_{2n+1}=1-c+b$, $w_{2n+1^*}=(1+s)(1-c)+b$ and $w_{2n^*}=1+s$ were obtained by estimating it from growth curves data that were previously obtained by~\citet{Yona2012} using mono-culture growth experiments.
The raw growth curves data were not published before, but were used to produce figures 3C, 4A, and S2 in \citet{Yona2012}.
We used \texttt{Curveball}\footnote{https://curveball.yoavram.com}, a dedicated method for predicting results of competition experiments from growth curve data~\citep{Ram2019}. \texttt{Curveball} takes growth curves of two strains growing separately in mono-culture and predicts how they would grow in a mixed culture, that is, it predicts the results of a competition assay.
From these predictions, relative fitness values can be computed.

We used growth curves of $2n+1$ and $2n$ in \SI{39}{\celsius} to estimate $1-c+b$. We used the same prior for $w_{2n+1^*}$ and $w_{2n^*}$, assuming that fitness values shouldn't be differ too much. We also tested different priors, but this one (let's call it main prior model) was better in sense of WAIC, prediction plots, estimated parameters. %TODO provide formula of WAIC and the results. To show ppc plots also?. 
\\We tested uninformative uniform prior, $\mathit{Uniform}(1,6)$, for all $w_{2n+1}$, $w_{2n+1^*}$, $w_{2n^*}$ or only for $w_{2n+1^*}$, $w_{2n^*}$. In both scenarios the estimated by the model parameters of fitness were looks too high: for the first scenario the estimated medians were  $w_{2n+1}=2.767$, $w_{2n+1^*}=5.79$, $w_{2n^*}=5.827$ and for the second $w_{2n+1}=1.046$, $w_{2n+1^*}=4.642$, $w_{2n^*}=4.684$. We don't expect to have such high fitness values. Moreover the WAIC for the models with these priors are 0.57 and 1.9 respectively, while WAIC for the main prior model is 0.27. Also for the model with uninformative prior for all fitnesses the fixation time of 2n+1 is about 16 in average, that contradicts unpublished observations that were about 250-300.
\\Also we tried to use additional growth curves. We used growth curves of 2n* (\emph{refined} from \citep{Yona2012}) and 2n+1 in \SI{39}{\celsius} to estimate $w_{2n*}/w_{2n+1}$. The same prior was used for $w_{2n*}/w_{2n+1*}$. This model not perfect well with WAIC 2.8.
\\Another try was to use growth curves of $2n+1$ and $2n$ in \SI{30}{\celsius} to estimate $1-c$. This estimation assumes that the cost of aneuploidy is the same in \SI{39}{\celsius} and \SI{30}{\celsius}; this might be incorrect, but we only assumed this to generate a prior distribution for the fitness values. The prior for $b$ was taken the same as for $c$. This model didn't work at all: we didn't receive any parameter that the likelihood with it is greater than zero. %TODO it make sense because bla bla?

Because \texttt{Curveball} uses a maximum-likelihood approach to estimate model parameters, we were able to estimate a distribution of relative fitness values by sampling from a truncated multivariate normal distribution defined by the maximum-likelihood covariance matrix. Thus, we sampled 10,000 values for appropriate fitness, which we used as prior distributions. 
See Figures~\ref{fig:growth-curves-30deg} and~\ref{fig:growth-curves-39deg}.
%TODO ILIA please report here the point estimate and 95% CI for the three distributions, I only have for 1-c+b: 0.9863 (95% CI 0.9790 - 0.9939).
%TODO ILIA describe the priors for the mutation and aneuploidy rate.
% TODO also describe the prior for the model without aneuploidy - the mutation rate is wider here.

@@

%%%% Fig - Curveball 39deg %%%
% generated with growth_curves.ipynb
\begin{figure}[h]
    \centering
	\includegraphics[width=0.5\textwidth]{../figures/evo39_fitness_39deg.pdf}
    \caption{
    \textbf{Fitness estimation from \SI{39}{\celsius}.}
    } 
    \label{fig:growth-curves-39deg}
\end{figure}
